### 2021-01-23 paper/intro.tex
Egits

\item \emph{syntactic} methods used in hardware models
  such as ARM or x86-TSO \cite{alglave},
\item \emph{alternate universe} methods (which give a semantics based on multiple speculative executions
  of the same program) such as the Java Memory Model~\cite{Manson:2005:JMM:1047659.1040336} 
  and related models \cite{Jagadeesan:2010:GOS:2175486.2175503, DBLP:conf/popl/KangHLVD17, DBLP:journals/pacmpl/ChakrabortyV19},
  % the speculative operational semantics of~\cite{Jagadeesan:2010:GOS:2175486.2175503},
  % the promising semantics of~\cite{DBLP:conf/popl/KangHLVD17},
  % or the event structures semantics of~\cite{DBLP:journals/pacmpl/ChakrabortyV19},
\item \emph{rewriting} methods, which give an operational model
  up to syntactic rewrites, such as~\cite{Pichon-Pharabod:2016:CSR:2837614.2837616}, and
\item \emph{logical} methods, such as the pomsets with preconditions
  model of~\cite{DBLP:journals/pacmpl/JagadeesanJR20}.

\end{itemize}
In this paper, we will focus on logical models, as those are compositional,
and align well with existing models of sequential composition.
The heart of the model of~\cite{DBLP:journals/pacmpl/JagadeesanJR20} is to add logical preconditions
to events, which are introduced by store actions (modeling data dependencies)
and conditionals (modeling control dependencies):
  \begin{gather*}
    \IF{s{<}1} \THEN z\GETS r\&s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wz0}{(s {<} 1) \land (r\&s){=}0 \mid \DW{z}{0}}{}
      \end{tikzinline}}
  \end{gather*}
Preconditions are discharged by being ordered after a read:
  \begin{gather*}
    r\GETS x\SEMI s\GETS y\SEMI \IF{s{<}1} \THEN z\GETS r\&s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{rx0}{\DR{x}{0}}{}
        \event{ry0}{\DR{y}{0}}{right=of rx0}
        \event{wz0}{(s{=}0) \limplies (s {<} 1) \land (r\&s){=}0 \mid \DW{z}{0}}{right=of ry0}
        \po{ry0}{wz0}
      \end{tikzinline}}
  \end{gather*}
Note that there is dependency order from $(\DR{y}{0})$ to $(\DW{z}{0})$
so the precondition for $(\DW{z}{0})$ only has to be satisfied assuming the hypothesis
$(s{=}0)$. There is no matching order from $(\DR{x}{0})$ to $(\DW{z}{0})$
which is why we do not assume the hypothesis $(r{=}0)$. The precondition on
$(\DW{z}{0})$ is a tautology, and so can be elided in the diagram:
  \begin{gather*}
    r\GETS x\SEMI s\GETS y\SEMI \IF{s{<}1} \THEN z\GETS r\&s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{rx0}{\DR{x}{0}}{}
        \event{ry0}{\DR{y}{0}}{right=of rx0}
        \event{wz0}{\DW{z}{0}}{right=of ry0}
        \po{ry0}{wz0}
      \end{tikzinline}}
  \end{gather*}
While existing models of relaxed memory have detailed treatmenmts of parallel composition,
they often give sequential composition little attention, either ignoring it altogether,
or treating it operationally with its usual small-step semantics. This paper
investigates how existing models of sequential composition interact with relaxed memory.

>>>>

\item \emph{syntactic} methods used in hardware models
  such as ARM or x86-TSO \cite{alglave},
\item \emph{speculative execution} methods (which give a semantics based on multiple executions
  of the same program) such as the Java Memory Model~\cite{Manson:2005:JMM:1047659.1040336} 
  and related models \cite{Jagadeesan:2010:GOS:2175486.2175503, DBLP:conf/popl/KangHLVD17, DBLP:journals/pacmpl/ChakrabortyV19},
  % the speculative operational semantics of~\cite{Jagadeesan:2010:GOS:2175486.2175503},
  % the promising semantics of~\cite{DBLP:conf/popl/KangHLVD17},
  % or the event structures semantics of~\cite{DBLP:journals/pacmpl/ChakrabortyV19},
\item \emph{rewriting} methods, which give an operational model
  up to syntactic rewrites, such as~\cite{Pichon-Pharabod:2016:CSR:2837614.2837616}, and
\item \emph{logical} methods, such as the pomsets with preconditions
  model of~\cite{DBLP:journals/pacmpl/JagadeesanJR20}.

\end{itemize}
In this paper, we will focus on logical models, as those are compositional,
and align well with existing models of sequential composition.
The heart of the model of~\cite{DBLP:journals/pacmpl/JagadeesanJR20} is to add logical preconditions
to events, which are introduced by store actions (modeling data dependencies)
and conditionals (modeling control dependencies):
  \begin{gather*}
    \IF{s{<}1} \THEN z\GETS r{*}s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wz0}{(s {<} 1) \land (r{*}s){=}0 \mid \DW{z}{0}}{}
      \end{tikzinline}}
  \end{gather*}
Preconditions are discharged by being ordered after a read:
  \begin{gather*}
    r\GETS x\SEMI s\GETS y\SEMI \IF{s{<}1} \THEN z\GETS r{*}s \FI
    \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{rx0}{\DR{x}{0}}{}
        \event{ry0}{\DR{y}{0}}{right=of rx0}
        \event{wz0}{(s{=}0) \limplies (s {<} 1) \land (r{*}s){=}0 \mid \DW{z}{0}}{right=of ry0}
        \po{ry0}{wz0}
      \end{tikzinline}}
  \end{gather*}
Note that there is dependency order from $(\DR{y}{0})$ to $(\DW{z}{0})$
so the precondition for $(\DW{z}{0})$ only has to be satisfied assuming the hypothesis
$(s{=}0)$. There is no matching order from $(\DR{x}{0})$ to $(\DW{z}{0})$
which is why we do not assume the hypothesis $(r{=}0)$. Nonetheless, the precondition on
$(\DW{z}{0})$ is a tautology, and so can be elided in the diagram:
  \begin{gather*}
    % r\GETS x\SEMI s\GETS y\SEMI \IF{s{<}1} \THEN z\GETS r{*}s \FI
    % \\
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{rx0}{\DR{x}{0}}{}
        \event{ry0}{\DR{y}{0}}{right=of rx0}
        \event{wz0}{\DW{z}{0}}{right=of ry0}
        \po{ry0}{wz0}
      \end{tikzinline}}
  \end{gather*}

While existing models of relaxed memory have detailed treatments of parallel composition,
they often give sequential composition little attention, either ignoring it altogether,
or treating it operationally with its usual small-step semantics. This paper
investigates how existing models of sequential composition interact with relaxed memory.
